What We Eval in the Shadows

### 2021-08-12 paper/main.tex
Update 6.obfus
Add 6.nss

\paragraph{Obfuscation}
Some packages use \eval to bypass the restrictions imposed by the \k{R CMD
  CHECK} tool. This tool enforces some well-formedness rules on a submitted
package before accepting it for inclusion in CRAN. Some static checks ensure
that the package code does not use certain restricted functions. Package authors
use \eval to obfuscate their code. The example mutates the variable \k{s} in
the package environment. However, that is locked by default and
\k{unlockBinding}, is a restricted function. To work around this, the call to
unlock is done through \eval.

>>>>

\paragraph{Non-standard Scoping}

\Eval allows functions to resolve symbols in different environments that the
current one. For example, the \k{data.table} redefines the subsetting
operator\footnote{It defines it as a triplet \texttt{[i, j, k]}, where expression
  \texttt{j} is used to calculate a subset of rows from \texttt{i} grouped by \texttt{k}.}
for data frames to enable compact queries. The following  
computes the average departure delays on 2014 Delta flights for each
origin airport.

\begin{minipage}{.95\textwidth}
\medskip
\begin{lstlisting}
> flights[carrier=="DL" & year == 2014, mean(dep_delay), by=origin]
#    origin       V1
# 1:    LGA 11.35083
# 2:    EWR 14.96171
# 3:    JFK 12.45414
\end{lstlisting}\medskip
\end{minipage}

At its core, the \k{[} operator uses \k{substitute} to capture its operand
expressions, transforms them, and evaluates them using \eval in the given data
frame.\footnote{We omit the code as the implementation spans over 1700 lines,\\
  \indent \cf
\url{https://github.com/Rdatatable/data.table/blob/master/R/data.table.R}} The
symbols \k{carrier}, \k{year}, \k{dep_delay} and \k{origin} are not looked up
in the current environment, but instead in the \k{flights} data frame.

The non-standard scoping is an important use case in R. It allows ergonomic
functions for interactive data exploration. The ergonomics come from the
reduced typing and syntactic noise at the cost of some ambiguity. Since R is
used primarily interactively, this trade-off is justifiable.

\paragraph{Obfuscation}
A few packages use \eval to bypass the restrictions imposed by the \k{R CMD
  CHECK} tool. This tool enforces some well-formedness rules on a submitted
package before accepting it for inclusion in CRAN. Some static checks ensure
that the package code does not use certain restricted functions. Package authors
use \eval to obfuscate their code. The example mutates the variable \k{s} in
the package environment. However, that is locked by default and
\k{unlockBinding}, is a restricted function. To work around this, the call to
unlock is done through \eval.
